---
title: "Untitled"
author: "Fernando Silva Aguilar"
date: "June 27, 2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Input of Segments file IBDLD
The file was transpose using unix. We read each file separately.

```{r, message=FALSE}
library(tidyverse)
library(data.table)
library(reshape2) 
library(gtools)
library(plyr)
library(cowplot)
library(lemon)
library(ggthemes)
library(patchwork)
rm(list = ls())

setwd("C:/Users/deer1/Google Drive/Alejandro/02. IBDLD/")
for (i in 1:10) {
  assign(paste0("seg",i),
         as.data.frame(fread(
paste0("GIBDLD_",i,".segment.transosed.txt.gz"),
header = F)))
}

```

## Manipulating the segment file
To each file we create the column name as the concatenated row 2 and 4 (iid1 and iid2).

```{r}
colnames(seg1) <- apply(seg1[c(2,4),], 2, function(x) paste(x[1],x[2],sep = "_"))
colnames(seg2) <- apply(seg2[c(2,4),], 2, function(x) paste(x[1],x[2],sep = "_"))
colnames(seg3) <- apply(seg3[c(2,4),], 2, function(x) paste(x[1],x[2],sep = "_"))
colnames(seg4) <- apply(seg4[c(2,4),], 2, function(x) paste(x[1],x[2],sep = "_"))
colnames(seg5) <- apply(seg5[c(2,4),], 2, function(x) paste(x[1],x[2],sep = "_"))
colnames(seg6) <- apply(seg6[c(2,4),], 2, function(x) paste(x[1],x[2],sep = "_"))
colnames(seg7) <- apply(seg7[c(2,4),], 2, function(x) paste(x[1],x[2],sep = "_"))
colnames(seg8) <- apply(seg8[c(2,4),], 2, function(x) paste(x[1],x[2],sep = "_"))
colnames(seg9) <- apply(seg9[c(2,4),], 2, function(x) paste(x[1],x[2],sep = "_"))
colnames(seg10) <- apply(seg10[c(2,4),], 2, function(x) paste(x[1],x[2],sep = "_"))
```

*Remove the rows corresponding to fid1, iid1, fid2, and iid2 *
```{r}
seg1 <- seg1[-c(1:4,nrow(seg1)),]
seg2 <- seg2[-c(1:4,nrow(seg2)),]
seg3 <- seg3[-c(1:4,nrow(seg3)),]
seg4 <- seg4[-c(1:4,nrow(seg4)),]
seg5 <- seg5[-c(1:4,nrow(seg5)),]
seg6 <- seg6[-c(1:4,nrow(seg6)),]
seg7 <- seg7[-c(1:4,nrow(seg7)),]
seg8 <- seg8[-c(1:4,nrow(seg8)),]
seg9 <- seg9[-c(1:4,nrow(seg9)),]
seg10 <- seg10[-c(1:4,nrow(seg10)),]
```

*Create vector with the name for the rows. See IBDLD manual for the names*
```{r}
ro <- c("start_bp","end_bp","start_snp","end_snp","tot_snp","tract_len_Kb")

#Create column to register the name for each row*

seg1$variable <- NA
seg2$variable <- NA
seg3$variable <- NA
seg4$variable <- NA
seg5$variable <- NA
seg6$variable <- NA
seg7$variable <- NA
seg8$variable <- NA
seg9$variable <- NA
seg10$variable <- NA

seg1$variable[1] <- c("chr")
seg2$variable[1] <- c("chr")
seg3$variable[1] <- c("chr")
seg4$variable[1] <- c("chr")
seg5$variable[1] <- c("chr")
seg6$variable[1] <- c("chr")
seg7$variable[1] <- c("chr")
seg8$variable[1] <- c("chr")
seg9$variable[1] <- c("chr")
seg10$variable[1] <- c("chr")

seg1$variable[2:nrow(seg1)] <- rep(ro,(nrow(seg1)-1)/6) # The six is the numer of parameters in ro
seg2$variable[2:nrow(seg2)] <- rep(ro,(nrow(seg2)-1)/6)
seg3$variable[2:nrow(seg3)] <- rep(ro,(nrow(seg3)-1)/6)
seg4$variable[2:nrow(seg4)] <- rep(ro,(nrow(seg4)-1)/6)
seg5$variable[2:nrow(seg5)] <- rep(ro,(nrow(seg5)-1)/6)
seg6$variable[2:nrow(seg6)] <- rep(ro,(nrow(seg6)-1)/6)
seg7$variable[2:nrow(seg7)] <- rep(ro,(nrow(seg7)-1)/6)
seg8$variable[2:nrow(seg8)] <- rep(ro,(nrow(seg8)-1)/6)
seg9$variable[2:nrow(seg9)] <- rep(ro,(nrow(seg9)-1)/6)
seg10$variable[2:nrow(seg10)] <- rep(ro,(nrow(seg10)-1)/6)
```


*melt the data frames to creat a 4 column file*
```{r}

l <- mixedsort(ls()[grep("seg",ls())])
for (i in 1:length(l)) {
  d <- get(l[i])
  d <- melt(d, id.vars = c("variable") )
  colnames(d)[2] <- "Pair"
  d <- subset(d, d$value != "")
  d <- d[order(d[,"variable"]),]
  
  assign(paste0("sa",i),
         d)
}
library(stringr)
k <- mixedsort(ls()[grep("sa",ls())])
for (i in 1:length(k)) {
  d <- get(k[i])
  d1 <- subset(d, d$variable == "start_bp")
  colnames(d1)[3] <- "Start_bp"
  d2 <- subset(d, d$variable == "end_bp")
  colnames(d2)[3] <- "End_bp"
  d3 <- subset(d, d$variable == "start_snp")
  colnames(d3)[3] <- "Start_SNP"
  d4 <- subset(d, d$variable == "end_snp")
  colnames(d4)[3] <- "End_SNP"
  d5 <- subset(d, d$variable == "tot_snp")
  colnames(d5)[3] <- "Total_SNP"
  
  #df <- left_join(d1[,-1],d2[,-1]) %>% left_join(d3[,-1]) %>% left_join(d4[,-1]) %>% left_join(d5[,-1])
  df <- cbind(d1[,-1],d2[,-1], d3[,-1],d4[,-1],d5[,-1])
  df <- df[,-c(3,5,7,9)]
  df$iid1 <- substr(df$Pair,1,as.numeric(gregexpr(paste(c("_C","_P"),collapse = "|"),df$Pair))-1) 
  
  df$iid2 <- substr(df$Pair,as.numeric(gregexpr(paste(c("_C","_P"),collapse = "|"),df$Pair))+1,str_length(df$Pair))
  
  df$Chromosome <- i
  assign(paste0("sb",i),
         df)
}


j <- mixedsort(ls()[grep("sb",ls())])
list_df <- lapply(j, get)
dt <- rbindlist(list_df)
dt$Start_bp <- as.numeric(paste0(dt$Start_bp))
dt$End_bp <- as.numeric(paste0(dt$End_bp))
dt$Total_SNP <- as.numeric(paste0(dt$Total_SNP))


```


```{r}
#Read the filtered GBS data. Filer for Monomorphic SNPs
rm(list = ls()[!ls() %in% c("dt","path","path1")])
library(openxlsx); library(plyr); library(tidyverse);library(data.table)

path1 = "C:/Users/deer1/Google Drive/Alejandro/01. Filtrado/"
gbs <-fread(paste0(path1,"Polymorphic_GBS.hmp.txt"))

p <- gbs[,c("rs#","chrom","pos")] # Subseting GBS data

KAL <- paste0("P",1:15)
l <- mixedsort(unique(dt$iid1))
l <- l[-grep("P",l)]

d <- dt
d1 <- subset(dt, !dt$iid1 %in% KAL & dt$iid2 %in% KAL)

v <- mixedsort(as.character(unique(d1$Pair)))

```


```{r}

## For columns with 1 when the SNP was inside IBD and NA otherwise
st <- Sys.time()
for (i in 1:length(l)) {
  a <- l[i]
  da <- d1[d1$iid1 %in% a,]
  df <- as.data.frame(p)
  a <- as.character(unique(da$Pair))
for (j in 1:length(a)) {
  b <- subset(da, da$Pair == a[j])
df[[paste0(a[j])]] <- NA
for (ii in 1:nrow(b)) {
  to_range <- which(df$chrom == b$Chromosome[ii] &
                     df$pos >= b$Start_bp[ii] & 
                     df$pos <= b$End_bp[ii])
  
  df[to_range,paste0(a[j])] <- 1 
  #df[,paste0("Tot")] <- apply(df[4:length(colnames(df))], 1, function(x) sum(x, na.rm = T))
  }
}
  assign(paste0("df_",l[i]),
         df)
}

s <- ls()[ls() %in% ls()[grep("df_",ls())]]
length(s)

for (i in 1:length(s)) {
  dg <- get(s[i])
  colnames(dg)[4] <- "Pru"
  if(ncol(dg) > 4){
    dg$Totl <- as.vector( apply(dg[,4:ncol(dg)],1,function(x) sum(x, na.rm = T)))
  } else {dg$Totl <- dg$Pru} 
  
  colnames(dg)[4] <- colnames(get(s[i]))[4]
  assign(paste("df_",s[i]),
         dg) 
}


s <- ls()[ls() %in% ls()[grep("df_ df",ls())]]
length(s)

for (i in 1:length(s)) {
  dg <- get(s[i])
  na <- colnames(dg)[-c(1:3,ncol(dg))]
  
    for (ii in 1:length(na)) {
      if(ncol(dg) > 5){
    dg[,na[ii]] <- dg[ ,na[ii]]/dg[,ncol(dg)]
                       } else {
    dg[,na[ii]] <- dg[,na[ii]]
  } }
  
  assign(paste("prueba_",s[i]),
         dg)
}

s1 <- ls()[ls() %in% ls()[grep("prueba",ls())]]
length(s1)

list_df = lapply(ls()[ls() %in% s1], get)
fd <- join_all(list_df, by=c('rs#',"chrom","pos"), type='left')
path2 = "C:/Users/deer1/Google Drive/Alejandro/03. Resultados/"
#write.xlsx(fd, paste0(path2,"Tabla_Contribution.xlsx"))

EN <- Sys.time()
EN - st

rm(list = ls()[!ls() %in% c("dt","path","path1","fd")])
```


## Tabla de contribucion de cada Progenitor P1:P15
```{r}
fig <- data.frame(Sum = (apply(fd[,-c(1,2,3)], 2, function(x) sum(x, na.rm = T))))
dim(fig)
fig <- fig %>% rownames_to_column("Pair")
fig$iid1 <- substr(fig$Pair,1,as.numeric(gregexpr(paste(c("_C","_P"), collapse = "|"), fig$Pair))-1)

fig$iid2 <- substr(fig$Pair,as.numeric(gregexpr(paste(c("_C","_P"),collapse = "|"),fig$Pair))+1,str_length(fig$Pair))
fig$Ciclo <- substr(fig$iid1,1,as.numeric(gregexpr(paste(c("_D"), collapse = "|"), fig$iid1))-1)
fig <- subset(fig, fig$Pair %in% fig$Pair[-grep("Totl", 
                                            fig$Pair)] )
fig1 <- fig[,c("iid1","iid2","Sum")]
ta <- reshape(fig1, idvar = "iid1", timevar = "iid2", direction = "wide")
colnames(ta) <- gsub("Sum.","", colnames(ta))
ta <- ta[mixedsort(rownames(ta)),mixedsort(colnames(ta))]
ta <- ta %>% remove_rownames() %>% column_to_rownames("iid1")

ta1 <- 100*ta/nrow(fd) # the total number of SNP are 7993

ta1[is.na(ta1)] <- 0
ta1$`Non-IBD` <- apply(ta1,1, function(x) (100 - sum(x)))
path <- "C:/Users/deer1/Google Drive/Alejandro/03. Resultados/"
ta1 <- ta1 %>% rownames_to_column("DH") 
write.xlsx(ta1,paste0(path,"Figure_Structure.xlsx"))

```


```{r}
#########################STRUCTURE GRAPH
rm(list = ls()[!ls() %in% c("dt","fd","ta1","fig")])

path <- "C:/Users/deer1/Google Drive/Alejandro/03. Resultados/"

look <- openxlsx::read.xlsx(paste0(path, "Figure_Structure.xlsx"))
look$Ciclo <- substr(look$DH,1,as.numeric(gregexpr("_D",look$DH))-1)

d1 <- melt(look, id.vars = c("DH", "Ciclo"))
d1$Ciclo <- as.factor(paste0(d1$Ciclo))
d1$DH <- as.factor(paste0(d1$DH))

## Cree un subset para poder visualizar la grafica ya que cada ciclo tiene mas de 100 DH lines.
## En este ejemplo grafique solo 10 DH lines del ciclo 0 al crear una base de datos d2

d2 <- subset(d1,d1$Ciclo == "C0") # Extraigo solo C0
d2 <- subset(d2, d2$DH %in% unique(d2$DH)[1:10]) # Solo las 10 primeras DH lines

d2 %>% mutate(variable = fct_relevel(variable, paste0("P",2:15))) %>% # Esto es para darle orden a los progenitores
ggplot( aes(DH, value, fill = variable)) +
  geom_col(color = "gray", size = 0.1) +
    #facet_grid(~Ciclo, switch = "x", scales = "free", space = "free") + # Esto es si se quiere graficar por Ciclo
  theme_minimal() + 
  labs(x = "", 
       y = "Progenitor Contribution (%)") +
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_discrete(expand = expansion(add = 1), position = "top") +
  theme(panel.spacing.x = unit(0.1, "lines"),
    axis.text.x = element_text(angle = 90, hjust = 0, 
                               vjust = 0.5, face = 'bold', size = 12),
    axis.text.x.top = element_text(vjust = 0.5),
    axis.text.y = element_blank(),
    axis.title = element_text(size = 20, face = 'bold'),
    panel.grid = element_blank(),
    strip.text = element_text(size = 15, face = 'bold', vjust = -1),
    legend.title = element_blank(),
    legend.text = element_text(size = 15),
    legend.key.width = unit(1,"cm"),
  #  legend.key.height = unit(1,"cm"),
    legend.position = "bottom")+ 
  guides(fill = guide_legend(nrow = 2)) # Para crear la leyenda en dos filas
  #+ scale_fill_manual(values = c("#1B9E77","#D95F02","#7570B3","#E7298A","#66A61E","#E6AB02","#A6761D","#40004B","#666666"  )) # Tenes que incluir 15 colores diferentes

```

